---
title: "A.thaliana_Model_Tests"
author: "Jiao"
date: "2018年9月5日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


# **7. Differences of biomass along with urbanisation**

Model: "difference between average biomass of a certain genotype in the herbivory and control treatment" ~ urbanization
Test how the difference between herbivory and control changes with urbanization.

```{r}
# select working dataframe: "mass.geno.BU"(with means)

# calculate difference of average biomass between herbivory and control treatment
df <- data.frame(genotype = mass.geno.BU[mass.geno.BU$treatment == "C", "genotype"],
                   df.BG = mass.geno.BU[mass.geno.BU$treatment == "C", "BG.biomass"] - mass.geno.BU[mass.geno.BU$treatment == "H", "BG.biomass"],
                   df.AG = mass.geno.BU[mass.geno.BU$treatment == "C", "AG.biomass"] - mass.geno.BU[mass.geno.BU$treatment == "H", "AG.biomass"],
                   df.rs = mass.geno.BU[mass.geno.BU$treatment == "C", "ratio"] - mass.geno.BU[mass.geno.BU$treatment == "H", "ratio"])

# creat new working dataframe: merge "diff" and "BU"
df.BU <- merge(df, BU, by = "genotype")
View(df.BU)

# model differences with urbanisation
## at 50m scale
lm.dfBG.50 <- lm(df.BG ~ d50, data = df.BU)
summary(lm.dfBG.50)
with(df.BU, plot(d50, df.BG))   ##plot
abline(lm.dfBG.50)


## Root biomass
lm.dfBG.50 <- lm(df.BG ~ d50, data = df.BU)
lm.dfBG.100 <- lm(df.BG ~ d100, data = df.BU)
lm.dfBG.200 <- lm(df.BG ~ d200, data = df.BU)
lm.dfBG.400 <- lm(df.BG ~ d400, data = df.BU)
lm.dfBG.800 <- lm(df.BG ~ d800, data = df.BU)
lm.dfBG.1600 <- lm(df.BG ~ d1600, data = df.BU)
lm.dfBG.2400 <- lm(df.BG ~ d2400, data = df.BU)
lm.dfBG.3200 <- lm(df.BG ~ d3200, data = df.BU)

## shoot biomass
lm.dfAG.50 <- lm(df.AG ~ d50, data = df.BU)
lm.dfAG.100 <- lm(df.AG ~ d100, data = df.BU)
lm.dfAG.200 <- lm(df.AG ~ d200, data = df.BU)
lm.dfAG.400 <- lm(df.AG ~ d400, data = df.BU)
lm.dfAG.800 <- lm(df.AG ~ d800, data = df.BU)
lm.dfAG.1600 <- lm(df.AG ~ d1600, data = df.BU)
lm.dfAG.2400 <- lm(df.AG ~ d2400, data = df.BU)
lm.dfAG.3200 <- lm(df.AG ~ d3200, data = df.BU)

## Root/shoot ratio
lm.dfRS.50 <- lm(df.rs ~ d50, data = df.BU)
lm.dfRS.100 <- lm(df.rs ~ d100, data = df.BU)
lm.dfRS.200 <- lm(df.rs ~ d200, data = df.BU)
lm.dfRS.400 <- lm(df.rs ~ d400, data = df.BU)
lm.dfRS.800 <- lm(df.rs ~ d800, data = df.BU)
lm.dfRS.1600 <- lm(df.rs ~ d1600, data = df.BU)
lm.dfRS.2400 <- lm(df.rs ~ d2400, data = df.BU)
lm.dfRS.3200 <- lm(df.rs ~ d3200, data = df.BU)


## Model selection
AIC(lm.dfBG.50, lm.dfBG.100, lm.dfBG.200, lm.dfBG.400, lm.dfBG.800, lm.dfBG.1600, lm.dfBG.2400, lm.dfBG.3200)   ## model d800 has lowest AIC value.
AIC(lm.dfAG.50, lm.dfAG.100, lm.dfAG.200, lm.dfAG.400, lm.dfAG.800, lm.dfAG.1600, lm.dfAG.2400, lm.dfAG.3200)   ## model d800 has lowest AIC value.
AIC(lm.dfRS.50, lm.dfRS.100, lm.dfRS.200, lm.dfRS.400, lm.dfRS.800, lm.dfRS.1600, lm.dfRS.2400, lm.dfRS.3200)   ## model d200 has lowest AIC value.

## summarize best models
summary(lm.dfBG.800)
summary(lm.dfAG.800)
summary(lm.dfRS.200)
```

Results:
Although significant p-values of urbanisation are shown for shoot biomass (d800, p=0.04469) and root/shoot ratio (d200, p=0.0439), both models possess lower R2 (0.1805, 0.1821) and F values but larger residual standard errors compared to their means.

# **8. Difference of urban and rural genotypes**

Model: biomass ~ herbivory + genotype + herbivory x genotype + urbanization(categorical) + herbivory x urbanization(categorical)

# **8.1 at 50m scale**

```{r}
# select working dataframes: "biomass_HT", "BU"

# categorize as rural and urban groups
urban50 <- BU %>% top_n(9, d50)      ###select the most urban genotypes
urban50 <- cbind(habitat = "urban", urban50)      ###add a new column marked as "urban" habitat 

rural50 <- BU %>% top_n(-9, d50)     ###select the most rural genotypes
rural50 <- cbind(habitat = "rural", rural50)      ###add a new column marked as "rural" habitat 

ru50 <- rbind(urban50, rural50)

bm.BU.50 <- merge(biomass_HT, ru50, by = "genotype")
View(bm.BU.50)

# create model
## shoot biomass
mod.AG.ru50 <- lm(AG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.50)
summary(mod.AG.ru50)
anova(mod.AG.ru50)

## root biomass
mod.BG.ru50 <- lm(BG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.50)
summary(mod.BG.ru50)
anova(mod.BG.ru50)

## root/shoot ratio
mod.rs.ru50 <- lm(ratio ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.50)
summary(mod.rs.ru50)
anova(mod.rs.ru50)
```


# **8.2 at 100m scale**

```{r}
# categorize as rural and urban groups
urban100 <- BU %>% top_n(9, d100)      ###select the most urban genotypes
urban100 <- cbind(habitat = "urban", urban100)      ###add a new column marked as "urban" habitat 

rural100 <- BU %>% top_n(-9, d100)     ###select the most rural genotypes
rural100 <- cbind(habitat = "rural", rural100)      ###add a new column marked as "rural" habitat 

ru100 <- rbind(urban100, rural100)

bm.BU.100 <- merge(biomass_HT, ru100, by = "genotype")

# create model
## shoot biomass
mod.AG.ru100 <- lm(AG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.100)

## root biomass
mod.BG.ru100 <- lm(BG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.100)

## root/shoot ratio
mod.rs.ru100 <- lm(ratio ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.100)

```


# **8.3 at 200m scale**

```{r}
# categorize as rural and urban groups
urban200 <- BU %>% top_n(9, d200)      ###select the most urban genotypes
urban200 <- cbind(habitat = "urban", urban200)      ###add a new column marked as "urban" habitat 

rural200 <- BU %>% top_n(-9, d200)     ###select the most rural genotypes
rural200 <- cbind(habitat = "rural", rural200)      ###add a new column marked as "rural" habitat 

ru200 <- rbind(urban200, rural200)

bm.BU.200 <- merge(biomass_HT, ru200, by = "genotype")

# create model
## shoot biomass
mod.AG.ru200 <- lm(AG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.200)

## root biomass
mod.BG.ru200 <- lm(BG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.200)

## root/shoot ratio
mod.rs.ru200 <- lm(ratio ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.200)

```



# **8.4 at 400m scale**

```{r}
# categorize as rural and urban groups
urban400 <- BU %>% top_n(9, d400)      ###select the most urban genotypes
urban400 <- cbind(habitat = "urban", urban400)      ###add a new column marked as "urban" habitat 

rural400 <- BU %>% top_n(-9, d400)     ###select the most rural genotypes
rural400 <- cbind(habitat = "rural", rural400)      ###add a new column marked as "rural" habitat 

ru400 <- rbind(urban400, rural400)

bm.BU.400 <- merge(biomass_HT, ru400, by = "genotype")

# create model
## shoot biomass
mod.AG.ru400 <- lm(AG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.400)

## root biomass
mod.BG.ru400 <- lm(BG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.400)

## root/shoot ratio
mod.rs.ru400 <- lm(ratio ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.400)

```



# **8.5 at 800m scale**

```{r}
# categorize as rural and urban groups
urban800 <- BU %>% top_n(9, d800)      ###select the most urban genotypes
urban800 <- cbind(habitat = "urban", urban800)      ###add a new column marked as "urban" habitat 

rural800 <- BU %>% top_n(-9, d800)     ###select the most rural genotypes
rural800 <- cbind(habitat = "rural", rural800)      ###add a new column marked as "rural" habitat 

ru800 <- rbind(urban800, rural800)

bm.BU.800 <- merge(biomass_HT, ru800, by = "genotype")

# create model
## shoot biomass
mod.AG.ru800 <- lm(AG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.800)

## root biomass
mod.BG.ru800 <- lm(BG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.800)

## root/shoot ratio
mod.rs.ru800 <- lm(ratio ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.800)

```


# **8.6 at 1600m scale**

```{r}
# categorize as rural and urban groups
urban1600 <- BU %>% top_n(9, d1600)      ###select the most urban genotypes
urban1600 <- cbind(habitat = "urban", urban1600)      ###add a new column marked as "urban" habitat 

rural1600 <- BU %>% top_n(-9, d1600)     ###select the most rural genotypes
rural1600 <- cbind(habitat = "rural", rural1600)      ###add a new column marked as "rural" habitat 

ru1600 <- rbind(urban1600, rural1600)

bm.BU.1600 <- merge(biomass_HT, ru1600, by = "genotype")

# create model
## shoot biomass
mod.AG.ru1600 <- lm(AG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.1600)

## root biomass
mod.BG.ru1600 <- lm(BG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.1600)

## root/shoot ratio
mod.rs.ru1600 <- lm(ratio ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.1600)

```


# **8.7 at 2400m scale**

```{r}
# categorize as rural and urban groups
urban2400 <- BU %>% top_n(9, d2400)      ###select the most urban genotypes
urban2400 <- cbind(habitat = "urban", urban2400)      ###add a new column marked as "urban" habitat 

rural2400 <- BU %>% top_n(-9, d2400)     ###select the most rural genotypes
rural2400 <- cbind(habitat = "rural", rural2400)      ###add a new column marked as "rural" habitat 

ru2400 <- rbind(urban2400, rural2400)

bm.BU.2400 <- merge(biomass_HT, ru2400, by = "genotype")

# create model
## shoot biomass
mod.AG.ru2400 <- lm(AG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.2400)

## root biomass
mod.BG.ru2400 <- lm(BG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.2400)

## root/shoot ratio
mod.rs.ru2400 <- lm(ratio ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.2400)

```



# **8.8 at 3200m scale**

```{r}
# categorize as rural and urban groups
urban3200 <- BU %>% top_n(9, d3200)      ###select the most urban genotypes
urban3200 <- cbind(habitat = "urban", urban3200)      ###add a new column marked as "urban" habitat 

rural3200 <- BU %>% top_n(-9, d3200)     ###select the most rural genotypes
rural3200 <- cbind(habitat = "rural", rural3200)      ###add a new column marked as "rural" habitat 

ru3200 <- rbind(urban3200, rural3200)

bm.BU.3200 <- merge(biomass_HT, ru3200, by = "genotype")

# create model
## shoot biomass
mod.AG.ru3200 <- lm(AG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.3200)

## root biomass
mod.BG.ru3200 <- lm(BG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.3200)

## root/shoot ratio
mod.rs.ru3200 <- lm(ratio ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.3200)

```


# **8.9 Select the best models based on AIC**

```{r}
AIC(mod.AG.ru50, mod.AG.ru100, mod.AG.ru200, mod.AG.ru400, mod.AG.ru800, mod.AG.ru1600, mod.AG.ru2400, mod.AG.ru3200)
AIC(mod.BG.ru50, mod.BG.ru100, mod.BG.ru200, mod.BG.ru400, mod.BG.ru800, mod.BG.ru1600, mod.BG.ru2400, mod.BG.ru3200)
AIC(mod.rs.ru50, mod.rs.ru100, mod.rs.ru200, mod.rs.ru400, mod.rs.ru800, mod.rs.ru1600, mod.rs.ru2400, mod.rs.ru3200)


```

Results:
For shoot biomass, three models (1600m, 2400m and 3200m) have same lowest AIC value.
For root biomass, five models (200m, 800m, 1600m, 2400m and 3200m) have same lowest AIC value.
For root/shoot ratio, two models (200m and 800m) have same lowest AIC value.

```{r}
# summarize the best models
summary(mod.AG.ru3200)
summary(mod.BG.ru3200)
summary(mod.rs.ru200)
```

Results:
(1) For shoot biomass: Terms (treatment.C, habitat.urban) are not included in this model. There is remarkable difference among genotypes and difference in rural area. But the model yields low R2 (0.2495) and low F-value (12.9 on 5 and 174 DF) and larger Residual standard error (0.02752) compared to the mean biomass (0.0581).
(2) For root biomass: Same results as shoot biomass are shown. Terms (treatment.C, habitat.urban) are not included in this model. There is remarkable difference among genotypes and difference in rural area. But the model yields low R2 (0.2418) and low F-value (12.42 on 5 and 174 DF) and larger Residual standard error (0.008017) compared to the mean biomass (1.415e-02).
(3) For root/shoot ratio: Terms (treatment.C, habitat.urban) are not included in this model. Only difference in rural area is shown. But the model has very low R2 (0.0271).


# **8.10 Plot the best models**

```{r}
# Calculate mean and sd:
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE),
      se = sd(x[[col]])/sqrt(length(x)))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func, varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

View(bm.BU.1600)

## plots at 1600m scale for root and shoot biomass
sum.AG1600 <- data_summary(bm.BU.1600,
                        varname = "AG.biomass",
                        groupnames = c("habitat", "treatment"))

sum.BG1600 <- data_summary(bm.BU.1600,
                        varname = "BG.biomass",
                        groupnames = c("habitat", "treatment"))




p.modAG <- ggplot(sum.AG1600, aes(x = habitat, y = AG.biomass, fill = treatment)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = AG.biomass - sd, ymax = AG.biomass + sd), width=.2,
                position = position_dodge(.9)) +
  labs(title = "Aboveground shoot biomass", x = "Habitat", y = "Shoot mass (mg)")

p.modBG <- ggplot(sum.BG1600, aes(x = habitat, y = BG.biomass, fill = treatment)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = BG.biomass - sd, ymax = BG.biomass + sd), width=.2,
                position = position_dodge(.9)) +
  labs(title = "Belowground root biomass", x = "Habitat", y = "Root mass (mg)")


p.modAG
p.modBG

## plot at 200m scale for root/shoot ratio
sum.ratio200 <- data_summary(bm.BU.200,
                        varname = "ratio",
                        groupnames = c("habitat", "treatment"))

p.modratio <- ggplot(sum.ratio200, aes(x = habitat, y = ratio, fill = treatment)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = ratio - sd, ymax = ratio + sd), width=.2,
                position = position_dodge(.9)) +
  labs(title = "Root/shoot ratio", x = "Habitat", y = "root/shoot")
p.modratio




```


# **9. Plot biomass and built-up ratio grouped by habitat and treatment**

Plot all individual data.

```{r}
# categorize observations as "rural and urban groups" with treatment based on built-up ratio
## Have done in "chapter7" and retrieve these dataframes
bm.BU.50
bm.BU.100
bm.BU.200
bm.BU.400
bm.BU.800
bm.BU.1600
bm.BU.2400
bm.BU.3200



## plot them
mod.AG.ru50 <- lm(AG.biomass ~ treatment + genotype + treatment*genotype + habitat + treatment*habitat, data = bm.BU.50)
dp.AG50 <- ggplot(bm.BU.50, aes(x=d50, y=AG.biomass)) +
  geom_point() 

dp.AG50

ggplot(bm.BU.50, aes(x=d50, y=AG.biomass, shape=habitat, color=treatment)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE)
```

